#Analyzing FitBit Data
##Peer Assignment 1 - Reproducible Research
By Williams Hidalgo

##About

This was the first project for the Reproducible Research course. The purpose of this project is  answer a series of questions about data collected from a FitBit.

##Data

The data for this assignment was downloaded from the course web site:

-Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

-The variables included in this dataset are:

-**steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)

-**date**: The date on which the measurement was taken in YYYY-MM-DD format

-**interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing data
Load the data:
```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
```

##What is mean total number of steps taken per day?
I'm going to ignore the missing values in the dataset.

1.Make a histogram of the total number of steps taken each day
```{r}
steps.date<-aggregate(steps~date,data=activity, FUN=sum)
barplot(steps.date$steps, names.arg=steps.date$date, xlab="date", ylab="steps", main="Total number of steps taken each day")
```

2.Calculate and report the *mean* and *median* total number of steps taken per day
```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

##What is the average daily activity pattern?
1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps.interval<-aggregate(steps~interval,data=activity, FUN=mean)
plot(steps.interval, main="Average daily activity pattern", type="l")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

##Imputing missing values

1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity))
```

2.Devise a strategy for filling in all of the missing values in the dataset.

I'm going to use the mean for the 5-minute intervals.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity<-merge(activity, steps.interval, by="interval", suffixes=c("", ".y"))
nas<-is.na(activity$steps)
activity$steps[nas]<-activity$steps.y[nas]
activity<-activity[,c(1:3)]
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
steps.date<-aggregate(steps~date, data=activity, FUN=sum)
barplot(steps.date$steps, names.arg=steps.date$date, xlab="date", ylab="steps", main="Mean and median total number of steps taken per day")
mean(steps.date$steps)
median(steps.date$steps)
```

There's no significant diferent between mean and median values.

##Are there differences in activity patterns between weekdays and weekends?

Using the weekdays() function. Note that for language compatibility, the weekends (i.e. not weekdays()), I include "s?bado" and "domingo", witch are saturday and sunday in spanish.

1.Create a new factor variable in the dataset with two levels "weekdays" and "weekends" indicating whether a given date is a weekday or weekend day.
```{r}
daytype <- function(date) {
  if (weekdays(as.Date(date)) %in% c("domingo", "sunday", "s?bado", "saturday")) {
    "weekend"
  } else {
    "weekday"
  }
}
activity$daytype <- as.factor(sapply(activity$date, daytype))
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
par(mfrow=c(2, 1))
for (type in c("weekend", "weekday")) {
  steps.type<-aggregate(steps~interval, data=activity, subset=activity$daytype==type, FUN=mean)
  plot(steps.type, type="l", main=type)
}
```

That's it!